Price of Anarchy for Mean Field Games

### 2018-07-04 PoA.tex
h_{var}(t) fixed

In the following, we intend to simplify the explicit solutions (\ref{eq:SC_MFG_1}) and (\ref{eq:SC_MKV_1}) for the social costs in the LQEMFG and LQEMKV problems. First, we consider the last integral, where we drop the superscripts ${MFG}$ and ${MKV}$ to address both problems simultaneously. By using the Riccati equation associated with $\bar{\eta}_t$ and the equation $\dot{\bar{x}}_t = (a^x+a^{\bar{x}}+(a^y+a^{\bar{y}})\bar{\eta}_t)\bar{x}_t$ (see Appendix \ref{appendix_solving_fbsdes}), we notice that
\begin{equation*}
\begin{split}
	&\int_0^T \bar{\eta}_t^2 \bar{x}_t^2 dt\\
	=& - \frac{1}{c(b_2 + \bar{b}_2)} \left[ \int_0^T \dot{\bar{\eta}}_t \bar{x}_t^2 dt -  \int_0^T ( (b_1 + \bar{b}_1 - b^y - b^{\bar{y}}) \bar{\eta}_t - (b^x + b^{\bar{x}}) ) \bar{x}_t^2 dt \right]\\
	= & - \frac{1}{c(b_2 + \bar{b}_2)} \left[ \bar{\eta}_T\bar{x}_T^2-\bar{\eta}_0\bar{x}_0^2-2\int_0^T\bar{\eta}_t\bar{x}_t\dot{\bar{x}}_t dt -  \int_0^T ( (b_1 + \bar{b}_1 - b^y - b^{\bar{y}}) \bar{\eta}_t - (b^x + b^{\bar{x}}) ) \bar{x}_t^2 dt \right] \\
	= & 2 \int_0^T \bar{\eta}_t^2 \bar{x}_t^2  - \frac{1}{c(b_2 + \bar{b}_2)} \left[ \bar{\eta}_T \bar{x}_T^2 - \bar{\eta}_0 \bar{x}_0^2 - \int_0^T  ((b_1 + \bar{b}_1 + b^y + b^{\bar{y}} ) \bar{\eta}_t + (b^x + b^{\bar{x}}) ) \bar{x}_t^2 dt \right],
\end{split}
\end{equation*}
where we applied integration by parts for the second equality. Thus, if we denote
\begin{equation*}
	\lambda := \frac{c^{MFG}}{c^{MKV}} = \frac{b_2}{b_2 + \bar{b}_2}\cdot \frac{ r + \bar{r}(1-\bar{s})^2 }{r + \bar{r}(1-\bar{s})}
\end{equation*}
the equations (\ref{eq:SC_MFG_1}) and (\ref{eq:SC_MKV_1}) can be rewritten as
\begin{equation}
\begin{split}
	SC^{MFG} = & h_{var}(t) +  \frac{1}{2} \int_0^T \left[ \bar{b}_1 \lambda \bar{\eta}_t^{MFG} + (q+ \bar{q}(1-s)^2) - \lambda (q + \bar{q}(1-s)) \right] (\bar{x}_t^{MFG})^2 dt, \\
	& + \frac{1}{2} \lambda \left( \bar{\eta}^{MFG}_0 \bar{x}_0^2 - \bar{\eta}^{MFG}_T (\bar{x}_T^{MFG})^2 \right) + \frac{1}{2} (q_T+\bar{q}_T(1-s_T)^2)(\bar{x}_T^{MFG})^2\\
	SC^{MKV} = & h_{var}(t) + \frac{1}{2} \bar{\eta}^{MKV}_0 \bar{x}_0^2,
\end{split}
\label{eq:SC_MFG_MKV_2}
\end{equation}
where $ h_{var}(t): = \frac{1}{2}\int_0^T (q+\bar{q}+(r+\bar{r})(a\eta_t)^2)v_tdt + \frac{1}{2}(q_T+\bar{q}_T)v_T  $.

@@

\begin{corollary}
    Assuming (\ref{eq:assumptions}), if the initial condition $\xi$ is such that $\mathbb{E}(\xi)=0$, then $PoA=1$.
\end{corollary}
\begin{proof}
    By the explicit equation for $\bar{x}_t$ (see Appendix \ref{appendix_solving_fbsdes}), $\bar{x}_0=0$ implies $\bar{x}_t=0,\ \forall t \in [0,T]$. Therefore by Proposition \ref{prop:diff_SC}, $\Delta SC=0$. From the explicit formula for the variance, $h_{var}(t)>0$ and therefore by equation (\ref{eq:SC_MFG_MKV_2}), $SC^{MKV}>0$. Therefore, $PoA=1$.
\end{proof}

@@

\begin{proposition}
	Assuming (\ref{eq:assumptions}), the initial condition $\xi$ satisfies $\mathbb{E}(\xi)\neq 0$, and assuming $\bar{b}_1\neq 0$, then 
	\begin{equation*}
	\lim_{\bar{b}_2 \to 0} PoA =: PoA^{0,\bar{b}_2} > 1\qquad and \qquad\lim_{\bar{b}_2 \to \infty} PoA = 1.
	\end{equation*}
	\label{prop:b2bar}
\end{proposition}
\begin{proof}
	For the case $\bar{b}_2 \to 0$, we have $\displaystyle \lambda \to \lambda^{0,\bar{b}_2} := \frac{r + \bar{r}(1- \bar{s})^2}{r + \bar{r}(1-\bar{s})}$
	and 
	$$B \to B^{0,\bar{b}_2} := \frac{b_2^2}{r + \bar{r}(1-\bar{s})^2} \neq 0,\ C^u \to C^{0,\bar{b}_2}_u >0 ,\ D^u \to D^{0,\bar{b}_2}_u \geq 0.$$
	Thus, the existence and uniqueness for solutions of the Riccati equations (\ref{eq:riccati_ut}) and (\ref{eq:riccati_wt}) associated with $u_t$ and $w_t$, respectively, in the limiting region $\bar{b}_2 \to 0$ still hold. Let's denote the limiting solutions by 
	$u^{0,\bar{b}_2}_t, w^{0,\bar{b}_2}_t$. By the bounded convergence theorem, the mean function $\bar{x}_t^{MFG}$ has the limit:
   	$$\lim_{\bar{b}_2 \to 0} \bar{x}_t^{MFG} = \bar{x}_0 e^{\int_0^t (b_1 + \bar{b}_1 - B^{0,\bar{b}_2}\cdot u^{0,\bar{b}_2}_t) dt} =: \bar{x}^{0,\bar{b}_2,MFG}_t.$$ 
   	By our assumption on $\bar{b}_1$, we have $A^u \neq A^w$ and thus, $u^{0,\bar{b}_2}_t\neq w^{0,\bar{b}_2}_t$ on a set of positive Lebesgue measure.
   	Thus, by the bounded convergence theorem and the assumption on $\xi$, we deduce
	$$ \lim_{\bar{b}_2 \to 0} \Delta SC = \frac{1}{2} B^{0,\bar{b}_2} \int_0^T \left( (u_t^{0,\bar{b}_2} - w_t^{0,\bar{b}_2})\cdot \bar{x}_t^{0,\bar{b}_2,MFG} \right)^2 dt \  > 0. $$
	Meanwhile, $\eta_t$ does not depend on $\bar{b}_2$, and therefore the variance $v_t$ also does not depend on $\bar{b}_2$. By the explicit formula, $0<v_t<\infty$ for $t>0$ and $0 \leq w^{0,\bar{b}_2}_0 < \infty$, and thus,
	\begin{equation*}
    \begin{split}
	    \lim_{\bar{b}_2 \to 0}SC^{MKV} &= h_{var}(t) + w^{0,\bar{b}_2}_0 \bar{x}_0^2, \\
	    \lim_{\bar{b}_2 \to 0}SC^{MKV}&<\infty, \\
	    \lim_{\bar{b}_2 \to 0}SC^{MKV}&>0 .       
    \end{split}
	\end{equation*}
	Hence, we deduce:
	$$ \lim_{\bar{b}_2 \to 0} PoA =: PoA^{0,\bar{b}_2} > 1. $$

	Now, for the case when $\bar{b}_2 \to \infty$,  we have $\lambda \to 0$
	and $B \to \infty,\ C^u \to 0,\ D^u \to 0$. Moreover, if we consider the magnitude w.r.t. $\bar{b}_2$, we have:
	$$ \lambda = \mathcal{O}(\bar{b}^{-1}_2),\ C^u = \mathcal{O}(\bar{b}_2^{-1}),\ D^u = \mathcal{O}(\bar{b}_2^{-1}),\ B = \mathcal{O}(\bar{b}_2^2),$$
	and 
	$$ \frac{\delta^\pm_u}{\sqrt{b_2 + \bar{b}_2}} \xrightarrow[\bar{b}_2 \to \infty]{} \pm \sqrt{ \frac{b_2 (q + \bar{q}(1-s))}{r + \bar{r}(1-\bar{s})} } =: \pm \delta^{\infty,\bar{b}_2}.$$
	Thus, from equation (\ref{eq:riccati_ut_sol}), for $t \in [0,T)$ we can deduce:
	\begin{equation*}
	\begin{split}
	 	(b_2 + \bar{b}_2)^{\frac{3}{2}}\cdot u_t &= \frac{ - \left( C^u\sqrt{b_2 + \bar{b}_2} + \zeta \frac{\delta^+_u}{\sqrt{b_2 + \bar{b}_2}}  \right) + \left( C^u\sqrt{b_2 + \bar{b}_2} + \zeta \frac{\delta^-_u}{\sqrt{b_2 + \bar{b}_2}}  \right) e^{- (\delta^+_u - \delta^-_u)(T-t)}}{ \left( \frac{\delta^-_u}{b_2 + \bar{b}_2} - \frac{b_2 (q_T + \bar{q}_T (1-s_T) )}{r + \bar{r}(1-\bar{s})} \right) - \left( \frac{\delta^+_u}{b_2 + \bar{b}_2} - \frac{b_2 (q_T + \bar{q}_T (1-s_T) )}{r + \bar{r}(1-\bar{s})} \right)e^{- (\delta^+_u - \delta^-_u)(T-t)} }\\
	 	& \xrightarrow[\bar{b}_2 \to \infty]{} \delta^{\infty, \bar{b}_2} (r + \bar{r}(1-\bar{s})^2) 
	 \end{split} 
	\end{equation*}
	where $\zeta = \frac{b_2 (r + \bar{r}(1-\bar{s})^2) }{r+ \bar{r}(1-\bar{s})} (q_T + \bar{q}_T(1-s_T))$. 
	Thus, we have $B \cdot u_t = \mathcal{O}\left({\bar{b}_2}^{1/2}\right)$ for $t \in [0,T)$ and 
	$$ \bar{x}_t^{MFG} = \bar{x}_0 e^{\int_0^t (b_1 + \bar{b}_1 - B\cdot u_s) ds} \xrightarrow[\bar{b}_2 \to \infty]{} 0,\ \forall t \in (0,T]. $$	
	As in the proof of Proposition \ref{prop:b2}, for $t \in [0,T)$:
	$$ (b_2 + \bar{b}_2) w_t \to \sqrt{(r + \bar{r}(1-\bar{s})^2) C^w}$$
	so that $w_t \to 0, \forall t \in [0,T)$ as $\bar{b}_2 \to \infty$. Hence, by the bounded convergence theorem, we deduce:
	$$ \lim_{\bar{b}_2 \to \infty} \Delta SC =  \frac{1}{2(r+ \bar{r}(1-\bar{s})^2)} \int_0^T \lim_{\bar{b}_2 \to \infty} \left( \frac{ (b_2 + \bar{b}_2)^{\frac{3}{2}} u_t }{\sqrt{b_2 + \bar{b}_2}} - (b_2 + \bar{b}_2)w_t \right)^2 \cdot \left(\bar{x}_t^{MFG} \right)^2 dt \to 0.
	$$	
	Moreover, as shown in the case $\bar{b}_2 \to 0$, $\eta_t$ as well as the variance $v_t$ are invariant w.r.t. $\bar{b}_2$ and $v_t>0$ for $t>0$. Thus,
	$$ \lim_{\bar{b}_2 \to \infty} SC^{MKV} = h_{var}(t)> 0.$$
	Therefore, we conclude that:
	$$ \lim_{\bar{b}_2 \to \infty} PoA = 1.$$
\end{proof}

>>>>

In the following, we intend to simplify the explicit solutions (\ref{eq:SC_MFG_1}) and (\ref{eq:SC_MKV_1}) for the social costs in the LQEMFG and LQEMKV problems. First, we consider the last integral, where we drop the superscripts ${MFG}$ and ${MKV}$ to address both problems simultaneously. By using the Riccati equation associated with $\bar{\eta}_t$ and the equation $\dot{\bar{x}}_t = (a^x+a^{\bar{x}}+(a^y+a^{\bar{y}})\bar{\eta}_t)\bar{x}_t$ (see Appendix \ref{appendix_solving_fbsdes}), we notice that
\begin{equation*}
\begin{split}
	&\int_0^T \bar{\eta}_t^2 \bar{x}_t^2 dt\\
	=& - \frac{1}{c(b_2 + \bar{b}_2)} \left[ \int_0^T \dot{\bar{\eta}}_t \bar{x}_t^2 dt -  \int_0^T ( (b_1 + \bar{b}_1 - b^y - b^{\bar{y}}) \bar{\eta}_t - (b^x + b^{\bar{x}}) ) \bar{x}_t^2 dt \right]\\
	= & - \frac{1}{c(b_2 + \bar{b}_2)} \left[ \bar{\eta}_T\bar{x}_T^2-\bar{\eta}_0\bar{x}_0^2-2\int_0^T\bar{\eta}_t\bar{x}_t\dot{\bar{x}}_t dt -  \int_0^T ( (b_1 + \bar{b}_1 - b^y - b^{\bar{y}}) \bar{\eta}_t - (b^x + b^{\bar{x}}) ) \bar{x}_t^2 dt \right] \\
	= & 2 \int_0^T \bar{\eta}_t^2 \bar{x}_t^2  - \frac{1}{c(b_2 + \bar{b}_2)} \left[ \bar{\eta}_T \bar{x}_T^2 - \bar{\eta}_0 \bar{x}_0^2 - \int_0^T  ((b_1 + \bar{b}_1 + b^y + b^{\bar{y}} ) \bar{\eta}_t + (b^x + b^{\bar{x}}) ) \bar{x}_t^2 dt \right],
\end{split}
\end{equation*}
where we applied integration by parts for the second equality. Thus, if we denote
\begin{equation*}
	\lambda := \frac{c^{MFG}}{c^{MKV}} = \frac{b_2}{b_2 + \bar{b}_2}\cdot \frac{ r + \bar{r}(1-\bar{s})^2 }{r + \bar{r}(1-\bar{s})}
\end{equation*}
the equations (\ref{eq:SC_MFG_1}) and (\ref{eq:SC_MKV_1}) can be rewritten as
\begin{equation}
\begin{split}
	SC^{MFG} = & h_{var} +  \frac{1}{2} \int_0^T \left[ \bar{b}_1 \lambda \bar{\eta}_t^{MFG} + (q+ \bar{q}(1-s)^2) - \lambda (q + \bar{q}(1-s)) \right] (\bar{x}_t^{MFG})^2 dt, \\
	& + \frac{1}{2} \lambda \left( \bar{\eta}^{MFG}_0 \bar{x}_0^2 - \bar{\eta}^{MFG}_T (\bar{x}_T^{MFG})^2 \right) + \frac{1}{2} (q_T+\bar{q}_T(1-s_T)^2)(\bar{x}_T^{MFG})^2\\
	SC^{MKV} = & h_{var} + \frac{1}{2} \bar{\eta}^{MKV}_0 \bar{x}_0^2,
\end{split}
\label{eq:SC_MFG_MKV_2}
\end{equation}
where $ h_{var}: = \frac{1}{2}\int_0^T (q+\bar{q}+(r+\bar{r})(a\eta_t)^2)v_tdt + \frac{1}{2}(q_T+\bar{q}_T)v_T  $.

@@

\begin{corollary}
    Assuming (\ref{eq:assumptions}), if the initial condition $\xi$ is such that $\mathbb{E}(\xi)=0$, then $PoA=1$.
\end{corollary}
\begin{proof}
    By the explicit equation for $\bar{x}_t$ (see Appendix \ref{appendix_solving_fbsdes}), $\bar{x}_0=0$ implies $\bar{x}_t=0,\ \forall t \in [0,T]$. Therefore by Proposition \ref{prop:diff_SC}, $\Delta SC=0$. From the explicit formula for the variance, $h_{var}>0$ and therefore by equation (\ref{eq:SC_MFG_MKV_2}), $SC^{MKV}>0$. Therefore, $PoA=1$.
\end{proof}

@@

\begin{proposition}
	Assuming (\ref{eq:assumptions}), the initial condition $\xi$ satisfies $\mathbb{E}(\xi)\neq 0$, and assuming $\bar{b}_1\neq 0$, then 
	\begin{equation*}
	\lim_{\bar{b}_2 \to 0} PoA =: PoA^{0,\bar{b}_2} > 1\qquad and \qquad\lim_{\bar{b}_2 \to \infty} PoA = 1.
	\end{equation*}
	\label{prop:b2bar}
\end{proposition}
\begin{proof}
	For the case $\bar{b}_2 \to 0$, we have $\displaystyle \lambda \to \lambda^{0,\bar{b}_2} := \frac{r + \bar{r}(1- \bar{s})^2}{r + \bar{r}(1-\bar{s})}$
	and 
	$$B \to B^{0,\bar{b}_2} := \frac{b_2^2}{r + \bar{r}(1-\bar{s})^2} \neq 0,\ C^u \to C^{0,\bar{b}_2}_u >0 ,\ D^u \to D^{0,\bar{b}_2}_u \geq 0.$$
	Thus, the existence and uniqueness for solutions of the Riccati equations (\ref{eq:riccati_ut}) and (\ref{eq:riccati_wt}) associated with $u_t$ and $w_t$, respectively, in the limiting region $\bar{b}_2 \to 0$ still hold. Let's denote the limiting solutions by 
	$u^{0,\bar{b}_2}_t, w^{0,\bar{b}_2}_t$. By the bounded convergence theorem, the mean function $\bar{x}_t^{MFG}$ has the limit:
   	$$\lim_{\bar{b}_2 \to 0} \bar{x}_t^{MFG} = \bar{x}_0 e^{\int_0^t (b_1 + \bar{b}_1 - B^{0,\bar{b}_2}\cdot u^{0,\bar{b}_2}_t) dt} =: \bar{x}^{0,\bar{b}_2,MFG}_t.$$ 
   	By our assumption on $\bar{b}_1$, we have $A^u \neq A^w$ and thus, $u^{0,\bar{b}_2}_t\neq w^{0,\bar{b}_2}_t$ on a set of positive Lebesgue measure.
   	Thus, by the bounded convergence theorem and the assumption on $\xi$, we deduce
	$$ \lim_{\bar{b}_2 \to 0} \Delta SC = \frac{1}{2} B^{0,\bar{b}_2} \int_0^T \left( (u_t^{0,\bar{b}_2} - w_t^{0,\bar{b}_2})\cdot \bar{x}_t^{0,\bar{b}_2,MFG} \right)^2 dt \  > 0. $$
	Meanwhile, $\eta_t$ does not depend on $\bar{b}_2$, and therefore the variance $v_t$ also does not depend on $\bar{b}_2$. By the explicit formula, $0<v_t<\infty$ for $t>0$ and $0 \leq w^{0,\bar{b}_2}_0 < \infty$, and thus,
	\begin{equation*}
    \begin{split}
	    \lim_{\bar{b}_2 \to 0}SC^{MKV} &= h_{var} + w^{0,\bar{b}_2}_0 \bar{x}_0^2, \\
	    \lim_{\bar{b}_2 \to 0}SC^{MKV}&<\infty, \\
	    \lim_{\bar{b}_2 \to 0}SC^{MKV}&>0 .       
    \end{split}
	\end{equation*}
	Hence, we deduce:
	$$ \lim_{\bar{b}_2 \to 0} PoA =: PoA^{0,\bar{b}_2} > 1. $$

	Now, for the case when $\bar{b}_2 \to \infty$,  we have $\lambda \to 0$
	and $B \to \infty,\ C^u \to 0,\ D^u \to 0$. Moreover, if we consider the magnitude w.r.t. $\bar{b}_2$, we have:
	$$ \lambda = \mathcal{O}(\bar{b}^{-1}_2),\ C^u = \mathcal{O}(\bar{b}_2^{-1}),\ D^u = \mathcal{O}(\bar{b}_2^{-1}),\ B = \mathcal{O}(\bar{b}_2^2),$$
	and 
	$$ \frac{\delta^\pm_u}{\sqrt{b_2 + \bar{b}_2}} \xrightarrow[\bar{b}_2 \to \infty]{} \pm \sqrt{ \frac{b_2 (q + \bar{q}(1-s))}{r + \bar{r}(1-\bar{s})} } =: \pm \delta^{\infty,\bar{b}_2}.$$
	Thus, from equation (\ref{eq:riccati_ut_sol}), for $t \in [0,T)$ we can deduce:
	\begin{equation*}
	\begin{split}
	 	(b_2 + \bar{b}_2)^{\frac{3}{2}}\cdot u_t &= \frac{ - \left( C^u\sqrt{b_2 + \bar{b}_2} + \zeta \frac{\delta^+_u}{\sqrt{b_2 + \bar{b}_2}}  \right) + \left( C^u\sqrt{b_2 + \bar{b}_2} + \zeta \frac{\delta^-_u}{\sqrt{b_2 + \bar{b}_2}}  \right) e^{- (\delta^+_u - \delta^-_u)(T-t)}}{ \left( \frac{\delta^-_u}{b_2 + \bar{b}_2} - \frac{b_2 (q_T + \bar{q}_T (1-s_T) )}{r + \bar{r}(1-\bar{s})} \right) - \left( \frac{\delta^+_u}{b_2 + \bar{b}_2} - \frac{b_2 (q_T + \bar{q}_T (1-s_T) )}{r + \bar{r}(1-\bar{s})} \right)e^{- (\delta^+_u - \delta^-_u)(T-t)} }\\
	 	& \xrightarrow[\bar{b}_2 \to \infty]{} \delta^{\infty, \bar{b}_2} (r + \bar{r}(1-\bar{s})^2) 
	 \end{split} 
	\end{equation*}
	where $\zeta = \frac{b_2 (r + \bar{r}(1-\bar{s})^2) }{r+ \bar{r}(1-\bar{s})} (q_T + \bar{q}_T(1-s_T))$. 
	Thus, we have $B \cdot u_t = \mathcal{O}\left({\bar{b}_2}^{1/2}\right)$ for $t \in [0,T)$ and 
	$$ \bar{x}_t^{MFG} = \bar{x}_0 e^{\int_0^t (b_1 + \bar{b}_1 - B\cdot u_s) ds} \xrightarrow[\bar{b}_2 \to \infty]{} 0,\ \forall t \in (0,T]. $$	
	As in the proof of Proposition \ref{prop:b2}, for $t \in [0,T)$:
	$$ (b_2 + \bar{b}_2) w_t \to \sqrt{(r + \bar{r}(1-\bar{s})^2) C^w}$$
	so that $w_t \to 0, \forall t \in [0,T)$ as $\bar{b}_2 \to \infty$. Hence, by the bounded convergence theorem, we deduce:
	$$ \lim_{\bar{b}_2 \to \infty} \Delta SC =  \frac{1}{2(r+ \bar{r}(1-\bar{s})^2)} \int_0^T \lim_{\bar{b}_2 \to \infty} \left( \frac{ (b_2 + \bar{b}_2)^{\frac{3}{2}} u_t }{\sqrt{b_2 + \bar{b}_2}} - (b_2 + \bar{b}_2)w_t \right)^2 \cdot \left(\bar{x}_t^{MFG} \right)^2 dt \to 0.
	$$	
	Moreover, as shown in the case $\bar{b}_2 \to 0$, $\eta_t$ as well as the variance $v_t$ are invariant w.r.t. $\bar{b}_2$ and $v_t>0$ for $t>0$. Thus,
	$$ \lim_{\bar{b}_2 \to \infty} SC^{MKV} = h_{var}> 0.$$
	Therefore, we conclude that:
	$$ \lim_{\bar{b}_2 \to \infty} PoA = 1.$$
\end{proof}
